---
title: 'Formulae'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
<<<<<<< HEAD
library(fontawesome)
```

<div class="boxy boxy-primary">
#### <abbr title="Variance tells you the degree of spread in your data set. The more spread the data, the larger the variance is in relation to the mean.">Variance</abbr>

$s^2 = \frac{\Sigma(X-\bar{X})^2}{N-1}$

`var(mean)`

</div>




<div class="boxy boxy-primary">
#### Standard Deviation

$s = \sqrt{\frac{\Sigma(X-\bar{X})^2}{N-1}}$

`sd(X)`
</div>

<div class="boxy boxy-primary">
#### 95% Confidence limits on z (upper and lower)

$X = \mu \pm 1.96\sigma$
</div>

<div class="boxy boxy-primary">
  #### <abbr title="Gives the Standard Deviation of a set of data as a standardised number">Z Score</abbr>

  ${z = \frac{X - \bar{X}}{\sigma}}$

  given an area, find the boundary value that determines this area.

```{r}
  qnorm(0.5,mean=100,sd=15)
```
</div>
=======
```

| Purpose       | Formula                                    | R Input                  |
|---------------|:------------------------------------------:|--------------------------|
| Variance      | $s^2 = \frac{\Sigma(X-\bar{X})}{N-1}$      | `var(X)`                 |
| Std Deviation | $s = \sqrt{\frac{\Sigma(X-\bar{X})}{N-1}}$ | `sd(x)`                  |
| z-score       | ${z = \frac{X - \bar{X}}{\sigma}}$         | See below |

| Purpose               | R Input                                                        |
|-----------------------|----------------------------------------------------------------|
| Density               | `dnorm(x, mean = 0, sd = 1, log = FALSE)`                      |
| Distribution Function | `pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)` |
| Quantile Function     | `qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)` |
| Random Deviates       | `rnorm(n, mean = 0, sd = 1)`                                   |
      
---

#### Confidence limits on z

$X = \mu \pm 1.96\sigma$

---

#### Chi-Square

Goodness of Fit / 1 Way Classification

$\chi^2 = \sum\frac{(O-E)^2}{E}$

$df = C - 1$

Independence Test / 2 Way Classification

$E_{ij} = \frac{R_{i}C_{j}}{N}$

$df = (R-1)(C-1)$

Sample Size

$\Phi = \sqrt{\frac{\chi^2}{N}}$

---

### Linear Relationships

#### Correlation

$r = \frac{cov_{xy}}{s_{x}s_{y}}$

$df = N-2$

---

#### Covariance

$\frac{(X-\bar{X})(Y-\bar{Y})}{N-1}$

---

#### Significance of r

$t = \frac{r\sqrt{N-2}}{\sqrt{(1-r^2)}}$

---

#### Regression


| Purpose          | Formula                        |
|------------------|--------------------------------|
| General Equation | $\hat{Y} = bX - a$             |
| Slope            | $r = \frac{cov_{xy}}{s_{x}^2}$ |
| Intercept        | $a = \bar{Y} - b\bar{X}$       |

---
>>>>>>> f70a4454cc90d2fe1db69db2c4a622b5cf67cdaa
