---
title: "Practice Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)

PracQuiz <- read.csv(file = 'data/PracQuiz.csv')
```

```{r}
PracQuiz %>%
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```


What percentage of participants smoke?

-[X]    39%
-[ ]    61%
-[ ]    20%
-[ ]    82%

Question 2
How many participants smoke and report being "Somewhat active"?

-[ ]    14
-[X]    29
-[ ]    75
-[ ]    46

Question 3
What is the average for the variable pulse1?

-[ ]    119.09
-[ ]    75.00
-[ ]    10.91
-[X]    75.17

Question 4
What is the standard deviation for the variable height?

-[ ]    175.89
-[ ]    89.53
-[ ]    176.00
-[X]    9.46

Question 5
What is the range for the variable pusle2?

-[X]    97
-[ ]    94
-[ ]    51
-[ ]    148

Question 6
Nala scores 60 on pulse2. What is Nala's z-score when calculated based on the entire group of participants?

-[ ]    -33.49
-[ ]    1.64
-[X]    -1.64
-[ ]    33.49

Question 7
Anna scores 78 on pulse1. Tom scores 72 on pulse1. Emily scores 120 on pulse2. Who scores the highest relative to the other participants in the study?

-[ ]    Tom
-[X]    Emily
-[ ]    The scores are equally high
-[ ]    Anna

Question 8
What percentage of the general population can be expected to be taller than 185 cm?

-[ ]    83.15%
-[ ]    33.15%
-[X]    16.85%
-[ ]    25.16%

Question 9
7.21% of people in the general population are taller than Jason. How tall is Jason?

-[ ]    160.45 cm
-[ ]    200.30 cm
-[X]    189.70 cm
-[ ]    175.22 cm