---
title: "AT1"
author: "Michael Alford"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jmv)
library(kableExtra)
```

Psychology Research Methods (HPS201/HPS771) Assignment
Task 1

Study 1 - Questions

**1.   Conduct the appropriate analysis to examine if young adults are significantly more likely to agree with the statement “It’s cool to be famous on Instagram” compared to teenagers. Attach the relevant jamovi output to your submission. Report the results of the analysis in APA format including all relevant information.**

```{r, echo=FALSE}
  Instagram <- read.csv("data/Instagram.csv")
  options(digits = 3)
  jmv::contTables(
    formula = ~ Group:Famous,
    data = Instagram,
    phiCra = TRUE)
```

A Chi-square test for independence was conducted to examine whether fame on Instagram was perceived as cool dependent on a participant's age. The analysis revealed a significant association between the two variables, $\chi^2$(_df_ = 1, _N_ = 80) = 4.05, $p < .05$, $V = .55$.

**2.   Based on the data collected, Dr T-Tock concludes that participants who spend 90 minutes or more on social media per day are significantly more likely to agree (than disagree) with the statement “It’s cool to be famous on Instagram”. Is Dr T-Tock correct? Justify your response and attach the relevant jamovi output to your submission.**

```{r}

```
